Analysis_manuscript_mash_GxE

### 2020-11-13 R/Analysis_manuscript_mash_GxE.Rmd
Add greenup mash analysis section to manuscript

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(bigsnpr)
library(rlist)
#library(mashr)
#install_github("alice-macqueen/switchgrassGWAS")
library(switchgrassGWAS)
```

@@

```{r}
j=1 # phe_mult: 1 is fl
i=1
k=1 # subpop 1 = Gulf   3 = G&M

@@

View(U_hyp_comb[[2]])
U_hyp <- c(U_hyp, U_hyp_comb)
```
So between the 'single phenotype' U_hyp and the combined U_hyp, that will give me 16 U_hyp for the five flowering phenotypes, and 8 U_hyp for the greenup phenotypes.

@@

```{r}
j=2 # phe_mult 1 = flowering 2 = greenup
k=3 # subpop 1 = Gulf   3 = G&M

@@

m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
                           U_hyp = U_hyp,
                           suffix = paste0(names(phe_mult)[j], "_phe_",
                                           names(subpop_v)[k], "_subpop"),
                           saveoutput = TRUE)
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(bigsnpr)
library(rlist)
#library(mashr)
#install_github("alice-macqueen/switchgrassGWAS")
library(switchgrassGWAS)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

@@

```{r}
j=2 # phe_mult: 1 is fl
i=1
k=3 # subpop 1 = Gulf_and_Midwest   3 = Gulf

@@

#View(U_hyp_comb[[2]])
U_hyp <- c(U_hyp, U_hyp_comb)
```
So between the 'single phenotype' U_hyp and the combined U_hyp, that will give me 16 U_hyp for the five flowering phenotypes, and 8 U_hyp for the greenup phenotypes.

@@

```{r}
j=2 # phe_mult 1 = flowering 2 = greenup
k=3 # subpop 3 = Gulf   1 = G&M

@@

U_ed <- readRDS(paste0("analysis/gwas/BLUPs/single_sites/Mash_data_driven_covariances_", numSNPs, "_", names(phe_mult)[j], "_phe_", names(subpop_v)[k],
                       "_subpop", ".rds"))
U_hyp <- readRDS(paste0("analysis/gwas/BLUPs/single_sites/Mash_data_driven_covariances_", numSNPs, "_", names(phe_mult)[j], "_phe_", names(subpop_v)[3],
                       "_subpop", ".rds"))

m_out2 <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
                           U_hyp = U_hyp, U_ed = U_ed,
                           suffix = paste0(names(phe_mult)[j], "_phe_",
                                           names(subpop_v)[k], "_subpop"),
                           saveoutput = TRUE)

@@

```{r}
subpop_v  # 1 is G&M  3 is Gulf
numSNPs <- 5000
m1 <- readRDS(paste0("../analysis/gwas/BLUPs/single_sites/Strong_Effects", numSNPs, "SNPs_", names(phe_mult)[j], "_phe_", names(subpop_v)[1],
                       "_subpop", ".rds"))
m3 <- readRDS(paste0("../analysis/gwas/BLUPs/single_sites/Strong_Effects", numSNPs, "SNPs_", names(phe_mult)[j], "_phe_", names(subpop_v)[3],
                       "_subpop", ".rds"))

mash_plot_Ulist(m1,1)
mash_plot_Ulist(m3,1)
mash_plot_covar(m1)
mash_plot_covar(m3)
m1$loglik
m3$loglik
length(get_significant_results(m3))
```

```{r}
m <- readRDS(paste0("../analysis/gwas/BLUPs/single_sites/Strong_Effects5000SNPs_greenup_2_phe_Gulf_and_Midwest_subpop_Gulf_Uhyp.rds"))

mash_plot_covar(m)
mash_plot_sig_by_condition(m)
mash_plot_manhattan_by_condition(m)
mash_plot_pairwise_sharing(m)
mash_plot_effects(m, n = 6)

switchgrassGWAS::get_U_by_mass(m)
```

```{r}
library(mashr)
mash_plot_covar <- function(m, saveoutput = FALSE, suffix = ""){
  df <- get_estimated_pi(m)
  df <- enframe(df, name = "Covariance Matrix", value = "Mass") %>%
    mutate(`Covariance Matrix` = str_replace(.data$`Covariance Matrix`,
                                             "(Stand_)??Bhat_?",
                                             "single_effect_"),
           `Covariance Matrix` = str_replace(.data$`Covariance Matrix`,
                                             "^null$", "no_effects")) %>%
    arrange(desc(.data$`Covariance Matrix`))
  df$`Covariance Matrix` <- factor(df$`Covariance Matrix`,
                                   levels = (df$`Covariance Matrix`))
  ggobject <- ggplot(df) +
    geom_bar(aes(x = .data$Mass, y = .data$`Covariance Matrix`),
             stat = "identity") +
    switchgrassGWAS::theme_oeco

  if(saveoutput == TRUE){
    if(!(str_sub(suffix, end = 1) %in% c("", "_"))){
      suffix <- paste0("_", suffix)
    }
    if(str_sub(suffix, end = 1) %in% c("")){
      suffix <- paste0("_", get_date_filename())
    }
    save_plot(paste0("Covariance_matrix_mass_plot", suffix,
                     ".png"), plot = ggobject, base_aspect_ratio = 1,
              base_height = 4)
  }
  return(list(covar_df = df, ggobject = ggobject))
}
```

# ---------------------

# Greenup mash results

#' @param posterior_weights Vector containing the posterior
#'     probability of each mixture component in Ulist for the data.
```{r}
length(get_significant_results(m))
get_significant_results(m)[1:10]
get_significant_results(m)[1:10]

str(m$posterior_weights)

n=7500
i = get_significant_results(m)[n]

# largest probability that this SNP loads on this covariance matrix
max(m$posterior_weights[i,])
which(m$posterior_weights[i,] == max(m$posterior_weights[i,]))

# second largest probability that this SNP loads on this covariance matrix
max( m$posterior_weights[i,][m$posterior_weights[i,]!=max(m$posterior_weights[i,])])
which(m$posterior_weights[i,] == max(m$posterior_weights[i,][m$posterior_weights[i,]!=max(m$posterior_weights[i,])]))

sum(m$posterior_weights[,1])

mash_plot_covar(m, saveoutput = TRUE)

colnames(m$posterior_weights)
Ulist <- switchgrassGWAS:::get_Ulist(m)
names(Ulist)

df <- mashr:::get_estimated_pi(m, dimension = "all")
```

Gulf
```{r}
inputdir <- file.path("../analysis/gwas/BLUPs/single_sites", "Gulf")
mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
mash_out
m <- read_rds(file.path(inputdir, mash_out[11]))
mash_plot_covar(m, saveoutput = TRUE)
get_U_by_mass(m)
mash_plot_Ulist(m, range = get_U_by_mass(m))
mash_plot_effects(m, n = 1, saveoutput = TRUE)
mash_plot_sig_by_condition(m, saveoutput = TRUE)
mash_plot_manhattan_by_condition(m, saveoutput = TRUE)
mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[1])) # sign
mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[2])) # sign & size
Gulf_greenup_gxe <- get_GxE(m)

mash_plot_pairwise_sharing(corrmatrix = Gulf_greenup_gxe$S_AP/Gulf_greenup_gxe$S_all_pairwise, saveoutput = TRUE)
mash_plot_pairwise_sharing(corrmatrix = Gulf_greenup_gxe$S_DS/Gulf_greenup_gxe$S_all_pairwise, reorder = TRUE, saveoutput = TRUE)
mash_plot_pairwise_sharing(corrmatrix = Gulf_greenup_gxe$S_2_no, reorder = FALSE)
mash_plot_pairwise_sharing(corrmatrix = Gulf_greenup_gxe$S_all_pairwise, reorder = FALSE)

```{r}
821+2309+789+1401+1090+1277+658+1075
9420/10581
```

# SVD of a covariance matrix

Work through an two-condition example here to help me understand how to explain this.

Sample points from a bivariate Gaussian distribution with a standard deviation of 3 in roughly the lower left-upper right direction and of 1 in the orthogonal direction. Because the x and y components co-vary, the variances of x and y do not fully describe the distribution. A 2 × 2 covariance matrix is needed; the directions of the arrows correspond to the eigenvectors of this covariance matrix and their lengths to the square roots of the eigenvalues.
```{r}
cov1 <- matrix(c(1, -0.4, 0.3, -0.4, 1, 0.7, 0.3, 0.7, 1), ncol = 3 )

b1 <- svd(cov1)

b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 80%
b1$d[2] / sum(b1$d) # 20%
b1$d[3] / sum(b1$d) # 20%
# b1$u symmetric with v if first matrix is symmetric
b1$v
b1$d

#sum(b1$d[1:2]) / sum(b1$d) # First 2 eigenvectors explain 100% of total variance.
for(g in 1:3){
barplot(b1$v[,g]/cov1[which.max(abs(cov1[,g])),g])
}

L = chol(cov1)
nvars = dim(L)[1]
nobs=200
# Random variables that follow an M correlation matrix
r = t(L) %*% matrix(rnorm(nvars*nobs), nrow=nvars, ncol=nobs)
r = t(r)

rdata = as.data.frame(r)
names(rdata) = c('cond1', 'cond2', 'cond3')
library(lattice)
splom(rdata)
```
The closer the covariance between x and y is to zero, the more equal the eigenvalues will be, and the more equal the proportion of variance explained by each eigenvector will be.
If the covariance is negative the first eigenvector will show a sign change.

# ------------------
# Gulf & Midwest
```{r}
inputdir <- file.path("../analysis/gwas/BLUPs/single_sites", "Gulf_and_Midwest")
mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
mash_out
m <- read_rds(file.path(inputdir, mash_out[13]))

mash_plot_covar(m)
get_U_by_mass(m)
mash_plot_Ulist(m, range = get_U_by_mass(m))
mash_plot_effects(m, n = 1)
mash_plot_sig_by_condition(m)
mash_plot_manhattan_by_condition(m)
mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[1])) # sign
mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[2])) # sign & size
Gulf_greenup_gxe <- get_GxE(m)
```

```{r}
get_U_by_mass(m)
i=2
names(Ulist)[i]
Ulist <- switchgrassGWAS:::get_Ulist(m)
b1 <- svd(Ulist[[i]])
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
b1$d[2] / sum(b1$d) # get to 60%
b1$d[3] / sum(b1$d) # 
b1$d[4] / sum(b1$d) # 
b1$d[5] / sum(b1$d) #

sum(b1$d[1:2]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
for(g in 1:5){
barplot(b1$v[,g]/Ulist[[1]][which.max(abs(Ulist[[1]][,g])),g])
}

ggb1 <- tibble(`First eigenvector (PVE = 100%)` = b1$v[,1]/Ulist[[1]][which.max(abs(Ulist[[1]][,1])),1],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
  mutate(Function = str_remove(Function, "_$"),
         Garden = str_remove(Garden, "^_"))
ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
ggb1_plot <- ggb1 %>%
  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 100%)`)) +
  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_color_viridis_d(option = "B", end = 0.7, 
                      direction = -1) +
  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], ".png"), base_height = 4, plot = ggb1_plot)

ggb2 <- tibble(`First eigenvector (PVE = 21.7%)` = b1$v[,2]/Ulist[[1]][which.max(abs(Ulist[[1]][,2])),2],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
  mutate(Function = str_remove(Function, "_$"),
         Garden = str_remove(Garden, "^_"))
ggb2$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
ggb2_plot <- ggb2 %>%
  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 21.7%)`)) +
  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_color_viridis_d(option = "B", end = 0.7, 
                      direction = -1) +
  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], "Eigenvector_2.png"), base_height = 4, plot = ggb2_plot)

```
GO enrichment amongst things that load onto different covariance matrices?
